Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-20 paper/reviewer_response_PNAS.tex

The experimental approach of deep mutational scanning has been applied to viral proteins before, including HIV env, influenza H1 HA, and NP. The novelty in this paper is that it is the first time such an analysis has been performed for H3 specifically and that it demonstrates the (albeit limited) utility of deep mutational scanning in predicting the success of viral strains in nature. Ultimately, it is clear from the discussion that the authors see this study as an intermediate step, suggesting that their work is "an improvement over no information at all." While something is indeed better than nothing, in my opinion, this paper represents only an incremental step forward in the field of influenza evolutionary forecasting due to the aforementioned limitations, especially since this experiment and subsequent analysis could be performed using more stringent selection methods during viral passaging like treatment with human sera for data that is presumably more faithful to nature.

\response{This is a fair assessment by the reviewer.
However, this is the first paper to show that data from deep mutational scanning can actually provide some insight into evolutionary outcomes in nature, and it is possible that this can provide a useful framework for applying deep mutational scanning studies in the context of evolution in nature...
}
